---
title: "Lab1"
author: "Alyssa Winchell"
date: "January 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Objectives:
- basic review (data wrangkin)
-intro to spatial data in R with sf
-intro to shiny apps

```{r, include=FALSE}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)

```


2. Review: reading, wranglin, and visualizing data

a. Read in the data

```{r}

wine <- read_csv("wine_data.csv")

```

b. Wrangling

```{r}

wine_new<- wine %>% 
  select(country, province, winery, region_1, points, price) %>%  #select, columns; filter, row. can also use to pick desired ordering
  rename(state = province) %>% 
  filter(state == "California" | state == "Oregon" | state == "Washington") %>% 
  mutate(ppd = points/price) %>%  #to add new column while existing columns remain in this case
  arrange(-ppd) #by decreasing value

```

c. Use group_by and summarize() to find summary stats by group

```{r}

wine_summary <- wine_new %>% 
  group_by(state) %>% 
  summarize(
    mean_ppd = max(ppd, na.rm = TRUE), #MISSED WHAT THIS MEANS, FIND OUT, also whyyy unexpected symbol
    max_ppd = max(ppd, na.rm = TRUE))


```

d. A few gg-graphics

```{r}

wine_hists <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
  facet_wrap(~state, scales = "free") #separates groups into ind histos, free allows different scales

wine_hists

```

new graph: scatterplot of price vs points, with color determined by state, adding a linear model trendline on top of it

```{r}

wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color = state), alpha = 0.1) + #aes treats things like variables
  geom_smooth(color = "black", method = "lm", size = 0.4, se = FALSE) +
  scale_color_manual(values = c("royalblue", "mediumorchid1", "orange")) + # c means combination of
  facet_wrap(~state) +
  theme_light()

wine_points

```

2. California countries (spatual data) using sf

```{r}

ca_counties <- st_read(ds = ".", layer = "CA_Counties_TIGER2016") #will become eventually spatio-temp data

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

ca_pop_inc <- read_csv("ca_pop_inc.csv") %>% 
  rename(NAME = COUNTY)

ca_df <- full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome)

#plot(ca_counties)

```


d. Use geom_sf to make a map in ggplot

```{r}

ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome))

ca_income

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>% 
  addTiles() %>% 
  addPolygons()

```

